---
title: "Google Data Analytics Capstone_v2"
author: "Ganesh Rajasekar"
date: "`r Sys.Date()`"
output: pdf_document
---

The following document links and merges the output for the Google Data Analytics Certificate Capstone Project. In this project, my imaginary healthcare company, Rajasekar Inc. , will examine the relationships between Cancer Surgery volume, Primary Care Utilization, ED Utilization, and Inpatient Utilization from 2020-2024.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initialize packages


```{r Initialize Packages, include=FALSE}
library(tidyverse)
library(data.table)

```

# Read in Data

## Cancer Surgeries by Hospital

```{r pressure, echo=FALSE}
setwd("C:/Users/ganes/Desktop/Google Capstone Project Github Track/Cancer Surgeries California By Hospital")

cancsurg<-fread("ca-oshpd-cancer-surgeries-gachospitals-20132024_odp.csv")

#keep years 2020-2024 for now (switch to 2023 based on primary care availability)
cancsurg_2<-cancsurg[Year%in%c(2020:2024)]
#cancsurg_alt<-cancsurg%>%filter(Year>=2020&Year<=2024)

```

## ED utilization
```{r}
setwd("C:/Users/ganes/Desktop/Google Capstone Project Github Track/Emergency Room Visits")

edvisits<-fread("encounters-by-facility_2012-2024_20250807.csv")

#keep entries from 2020-2024 only
edvisits_2<-edvisits[year%in%c(2020:2024)]
edvisits_alt<-edvisits%>%filter(year>=2020&year<=2024)
```


## Primary Care Utilization

```{r}
setwd("C:/Users/ganes/Desktop/Google Capstone Project Github Track/Annual Primary Care Utilization 2020-2024/cleaned data")

filestoread<-list.files(getwd())

pcutil_20<-readxl::read_xlsx(filestoread[1],sheet="Page 1-8")
pcutil_21<-readxl::read_xlsx(filestoread[2],sheet="Page 1-8")
pcutil_22<-readxl::read_xlsx(filestoread[3],sheet="Page 1-8")
pcutil_23<-readxl::read_xlsx(filestoread[4],sheet="Page 1-8")
```

